---
title: "Regression Analysis (cf. 4.2)"

output:
  html_notebook: default
---

This R Notebook is used to load, fit, and compare various linear models (lms).

In more detail, the following is done:
1. The data is loaded and prepared.
2. Data inspection.
3. The following lms can then be fitted:
  - Simple lms.
  - Multiple lms.
  - Multiple lms with interactions terms.
  - Time-lagged lms.
  
Analyses are enumerated. However, these numbers do not necessarily follow
table numbers in the paper, but built upon each other. In addition, please note 
that not every analysis is reflected in detail in a table in the paper but can 
also be part of a larger analysis or as a basis for general conclusions and thus, 
for your reference.

To use this NB, the data has to be processed and prepared with the VK utils
module (run_preprocessing.sh and run_calculations.sh)

Load packages needed.

```{r}
## First specify the packages of interest
packages = c("dplyr", "Hmisc", "stringr", "tidyverse", "lmtest", "dyn", "markdown")

## Now load or install&load all
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

1. Load and prepare data.

```{r Load Data, warning=FALSE}

# Specify which time slice should be loaded >> Options: 7,5,3,1
# Use a string, as in the pre-selected time slice of "7" da


time_slice = 5 # choose time slice needed
data_dir <- paste("./../../../data/metrics_percent_results/", time_slice, "days/countries", sep = "")

# Save current working directory
oldwd <- getwd()

# Set data directory and read in data
setwd(data_dir)
files  <- list.files(pattern = '.csv')
tables <- lapply(files, read.csv, header = TRUE,  stringsAsFactors = FALSE)
combined <- do.call(rbind, tables)

# Go back to usual working directory
setwd(oldwd)
```

Convert status denoting subcorpus affiliation into 0 and 1 factor, where control = 0 and free = 1. 

```{r message=FALSE}
library(dplyr)
combined$status <- as.factor(combined$status)
combined <- combined %>% 
    mutate(status = recode(status,
                      "control" = "0", 
                      "free" = "1"))
# remove De Facto
combined <- combined[!grepl("de_facto_states", combined$country),]
countries = unique(combined$country)
```

Inspect data.

```{r message=FALSE, warning=FALSE}
library(Hmisc)
describe(combined)
summary(combined$post)
```

# Analysis 1 (cf. Supplementary Data, §2, Table 3: Free/Control, Single Countries)

We include a range of commented tests and analyses that were carrried out.
To use them, please uncomment. Please note that for the sake of readability, 
the commented lines are removed from later analyses. 


For subset in vk_corpus:
  for each country in subset:
    y <- delta_cov ~ delta_rtsi
    
a. Free subset
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lmtest)

par(mfrow = c(2, 2))

# Initialize list of all country lms
analysis_1_free_res <- list()

for (elem in countries) {
  # Create free subcorpus
  free <-
    combined %>% filter(combined$country == elem, combined$status == 1)
  
  # Experiments with absolute values
  # free$rtsi <- round((free$rtsi * 100),0)
  # free$abs_posts <- round((free$abs_posts * 100),0)
  # analysis_1 <- lm(abs_posts ~ rtsi, data = free)
  
  # Experiments with log-transformed values
  # free$rtsi_pct = log(free$rtsi_pct+1)
  # free$post = log(free$post+1)
  
  # Experiments with taking the square-root of percent and absolute values
  # free$rtsi_pct = sqrt(pmax(0, free$rtsi_pct))
  # free$post = sqrt(pmax(0, free$post))
  
  
  # Analysis 1: DeltaCoverage ~ DeltaRTSI
  analysis_1 <- lm(post ~ rtsi_pct, data = free)
  print(elem)
  
  # Check whether residuals are autocorrelated or not
  # acf(analysis_1$residuals)
  # runs <- lawstat::runs.test(analysis_1$residuals)
  # print(runs)
  
  # Durbin-Watson Test (Auto correlation of disturbances)
  # dw_test <- lmtest::dwtest(analysis_1)
  # print(dw_test)
  
  # Test for Association/Correlation Between Paired Samples
  # cor_test <- cor.test(free$rtsi_pct, analysis_1$residuals)
  # print(cor_test)
  
  # Check mean of residuals, which should be near 0
  # check_mean_of_res = mean(analysis_1)
  # print(check_mean_of_res)
  
  print(summary(analysis_1))
  
  # Check confidence intervals
  # print(confint(analysis_1))
  
  # Visual inspection
  plot(analysis_1)
  
  # Add to all other countries
  analysis_1_free_res[[elem]] <- analysis_1
}
```

b. Control subset
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lmtest)

par(mfrow = c(2, 2))

# Initialize list of all country lms
analysis_1_control <- list()

for (elem in countries) {
  # Create control subcorpus
  control <-
    combined %>% filter(combined$country == elem, combined$status == 0)
  
  # Experiments with absolute values
  # control$rtsi <- round((control$rtsi * 100),0)
  # control$abs_posts <- round((control$abs_posts * 100),0)
  # analysis_1 <- lm(abs_posts ~ rtsi, data = control)
  
  # Experiments with log-transformed values
  # control$rtsi_pct = log(control$rtsi_pct+1)
  # control$post = log(control$post+1)
  
  # Experiments with taking the square-root of percent and absolute values
  # control$rtsi_pct = sqrt(pmax(0, control$rtsi_pct))
  # control$post = sqrt(pmax(0, control$post))
  
  
  # Analysis 1: DeltaCoverage ~ DeltaRTSI
  analysis_1 <- lm(post ~ rtsi_pct, data = control)
  print(elem)
  
  # Check whether residuals are autocorrelated or not
  # acf(analysis_1$residuals)
  # runs <- lawstat::runs.test(analysis_1$residuals)
  # print(runs)
  
  # Durbin-Watson Test (Auto correlation of disturbances)
  # dw_test <- lmtest::dwtest(analysis_1)
  # print(dw_test)
  
  # Test for Association/Correlation Between Paired Samples
  # cor_test <- cor.test(control$rtsi_pct, analysis_1$residuals)
  # print(cor_test)
  
  # Check mean of residuals, which should be near 0
  # check_mean_of_res = mean(analysis_1)
  # print(check_mean_of_res)
  
  print(summary(analysis_1))
  
  # Check confidence intervals
  # print(confint(analysis_1))
  
  # Visual inspection
  plot(analysis_1)
  
  # Add to all other countries
  analysis_1_control[[elem]] <- analysis_1
}
```

# Analysis 2 # Analysis 1 (cf. Supplementary Data, §2, Table 3: Both, Single Countries)

For each country in vk_corpus
  y <- delta_cov ~ delta_rtsi

```{r message=FALSE, warning=FALSE}
library(tidyverse)

par(mfrow=c(2,2))

analysis_2_b <- list()

for (elem in countries) {
  
  # Use both subsets together
  both <- combined %>% filter(combined$country == elem)
  print(elem)
  
  # Analysis 2: DeltaCoverage ~ DeltaRTSI
  analysis_2 <- lm(post ~ rtsi_pct, data = both)
  print(summary(analysis_2))
  plot(analysis_2)
  analysis_2_b[[elem]] <- analysis_2
}
```

# Analysis 3 (Baseline analysis for analyszing significance of factors "status" and "country".

For all_countries in vk_corpus
  y <- delta_cov ~ delta_rtsi

```{r message=FALSE, warning=FALSE}
library(tidyverse)

par(mfrow=c(2,2))

# Analysis 3: DeltaCoverage ~ DeltaRTSI
analysis_3 <- lm(post ~ rtsi_pct, data = combined)
print(summary(analysis_3))
plot(analysis_3)
```

# Analysis 4 (cf. Supplementary Material, §2, Table 4)

For each country in vk_corpus:
    y <- delta_cov ∼ delta_rtsi + status
    
```{r message=FALSE, warning=FALSE}
library(tidyverse)

par(mfrow=c(2,2))

analysis_4_b <- list()

for (elem in countries) {
  
  # Use both subsets together
  both <- combined %>% filter(combined$country == elem)
  print(elem)
  
  # Analysis 4: DeltaCoverage ~ DeltaRTSI + status
  analysis_4 <- lm(post ~ rtsi_pct + status, data = both)
  print(summary(analysis_4))
  plot(analysis_4)
  analysis_4_b[[elem]] <- analysis_4
}
```

# Analysis 5 (Analysis reg. significance of factor "status", not outlined in detail in table) 

For all_countries in vk_corpus:
  y <- delta_cov ∼ delta_rtsi + status

To inspect the data as for Analysis 1, add the corresponding lines.
For the sake of readability, they are spared in the following.

```{r message=FALSE, warning=FALSE}
library(tidyverse)

par(mfrow=c(2,2))

# Analysis 5: DeltaCoverage ~ DeltaRTSI + status
analysis_5 <- lm(post ~ rtsi_pct + status, data = combined)
print(summary(analysis_5))
plot(analysis_5)
```

# Analysis 6 (cf. Supplementary Material, §2, Table 1)

For all countries in subset:
    y <- delta_cov ∼ delta_rtsi + country

a. Free subset
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lmtest)

par(mfrow = c(2, 2))

# Load data as needed, using Africa as reference factor
free <- combined %>% filter(combined$status == 1)
free$country <- as.factor(free$country)
free$country <- relevel(free$country, ref = 1)

# Create free subcorpus
free <-
  combined %>% filter(combined$status == 1)

# Analysis 6: DeltaCoverage ~ DeltaRTSI
analysis_6 <- lm(post ~ rtsi_pct + country, data = free)

print(summary(analysis_6))

# Visual inspection
plot(analysis_6)
```

b. Control subset
```{r message=FALSE, warning=FALSE, cache=TRUE}
library(tidyverse)
library(lmtest)

par(mfrow = c(2, 2))

# Load data as needed, using Africa as reference factor
control <- combined %>% filter(combined$status == 0)
control$country <- as.factor(control$country)
control$country <- relevel(control$country, ref = 1)

# Create control subcorpus
control <-
combined %>% filter(combined$status == 0)

# Analysis 6: DeltaCoverage ~ DeltaRTSI
analysis_6 <- lm(post ~ rtsi_pct + country, data = control)

print(summary(analysis_6))

# Visual inspection
plot(analysis_6)
```

# Analysis 7 (Analysis reg. significance of factor "country", not outlined in detail in table) 

For all countries in vk_corpus:
  y <- delta_cov ∼ delta_rtsi + country

```{r message=FALSE, warning=FALSE, cache=TRUE}
library(tidyverse)

par(mfrow = c(2, 2))

# Load data as needed, using Africa as reference factor
combined$country <- as.factor(combined$country)
combined$country <- relevel(combined$country, ref = 1)

analysis_7 <- lm(post ~ rtsi_pct + country, data = combined)
summary(analysis_7)
plot(analysis_7)
```

# Analysis 8 (Analysis reg. significance of factor "country", not outlined in detail in table) 

For all countries in subset:
    y <- delta_cov ∼ delta_rtsi + country + (delta_rtsi * country)

a. Free subset
```{r message=FALSE, warning=FALSE, cache=TRUE}
library(tidyverse)
library(lmtest)

par(mfrow = c(2, 2))

# Load data as needed, using Africa as reference factor
free <- combined %>% filter(combined$status == 1)
free$country <- as.factor(free$country)
free$country <- relevel(free$country, ref = 1)

# Analysis 8: DeltaCoverage ~ DeltaRTSI + country + (DeltaRTSI * country)
analysis_8 <- lm(post ~ rtsi_pct + country + (rtsi_pct * country), data = free)
print(summary(analysis_8))
plot(analysis_8)
```

b. Control subset
```{r message=FALSE, warning=FALSE, cache=TRUE}
library(tidyverse)
library(lmtest)

par(mfrow = c(2, 2))

# Initialize list of all country lms
analysis_8_f <- list()

# Load data as needed, using Africa as reference factor
control <- combined %>% filter(combined$status == 0)
control$country <- as.factor(control$country)
control$country <- relevel(control$country, ref = 1)

# Analysis 8: DeltaCoverage ~ DeltaRTSI + country + (DeltaRTSI * country)
analysis_8 <- lm(post ~ rtsi_pct + country + (rtsi_pct * country), data = control)
summary(analysis_8)
plot(analysis_8)
```

# Analysis 9 (Analysis reg. significance of factor "country", not outlined in detail in table) 

for all_countries in vk_corpus:
  y <- delta_cov ∼ delta_rtsi + country + (delta_rtsi * country)
  
```{r message=FALSE, warning=FALSE, cache=TRUE}
library(tidyverse)

par(mfrow = c(2, 2))

# Load data as needed, using Africa as reference factor
combined$country <- as.factor(combined$country)
combined$country <- relevel(combined$country, ref = 1)

analysis_9 <- lm(post ~ rtsi_pct + country + (rtsi_pct * country), data = combined)
summary(analysis_9)
plot(analysis_9)
```

# Model Comparison II (Analysis reg. significance of factor "status" and "country", cf. §4.2) 

```{r message=FALSE, warning=FALSE, cache=TRUE}
# Model Comparison reg. significance of factors (status and country)

# Check significance of status
anova(analysis_3, analysis_5, test="Chisq") 
anova(analysis_3, analysis_7, analysis_9, test="Chisq")

# Check significance of country
anova(analysis_3, analysis_7, test="Chisq") 
anova(analysis_6, analysis_8, test="Chisq")
```

# Time-Lagged Linear Regression (cf. §4.2, Table 4)
# Analysis 10

y <- delta_cov ~ lag(delta_cov,-1) + lag(delta_rtsi,-1)

# Analysis 11

y <- delta_cov ~ lag(delta_cov,-1) + lag(delta_cov,-2) + lag(delta_rtsi,-1) +
lag(delta_cov,-2) + lag(delta_rtsi,-2)

10.a and 11.a Free subset
```{r message=FALSE, warning=FALSE, cache=TRUE}
library(tidyverse)
library(dyn)
library(zoo)
library(xts)
par(mfrow = c(2, 2))

analysis_10_f <- list()
analysis_11_f <- list()

combined$country <- as.factor(combined$country)
combined$country <- relevel(combined$country, ref = 1)

for (elem in countries) {
  
  # Get free subcorpus
  free <- combined %>% filter(combined$country == elem, combined$status == 1)
  # Choose the right lag from stats
  lag <- stats::lag
  # Create time series
  dates <- free$date
  dates <- seq(as.Date('2018-01-17'), length = 52, by = "days")
  free_sel <- data.frame(free$rtsi_pct, free$post)
  free_ts <- zoo(free_sel, dates)
  
  # Analysis 10.a: DeltaCoverage (t) ~ DeltaCoverage (t-1) + DeltaRTSI (t-1)
 analysis_10 <-
    dyn$lm(free.post ~ lag(free.post,-1) + lag(free.rtsi_pct,-1), data = free_ts)
  
  # Analysis 11.a: DeltaCoverage (t) ~ DeltaCoverage (t-2) + DeltaRTSI (t-2)
  analysis_11 <-
    dyn$lm(
      free.post ~ lag(free.post,-1) + lag(free.rtsi_pct,-1) + lag(free.post, -2) + lag(free.rtsi_pct, -2),
      data = free_ts
    )

  # Plot and inspect results
  print(elem)

  print(summary(analysis_10))
  print(summary(analysis_11))
  
  plot(analysis_10)
  plot(analysis_11)
  
  # Save results
  analysis_10_f[[elem]] <- analysis_10
  analysis_11_f[[elem]] <- analysis_11
}
```

```{r message=FALSE, warning=FALSE, cache=TRUE}
library(tidyverse)
library(dyn)
library(zoo)
library(xts)

par(mfrow = c(2, 2))

analysis_10_c <- list()
analysis_11_c <- list()

combined$country <- as.factor(combined$country)
combined$country <- relevel(combined$country, ref = 1)

for (elem in countries) {
  
  # Get control subcorpus
  control <- combined %>% filter(combined$country == elem, combined$status == 0)
  # Choose the right lag from stats
  lag <- stats::lag
  # Create time series
  dates <- control$date
  dates <- seq(as.Date('2018-01-17'), length = 52, by = "days")
  control_sel <- data.frame(control$rtsi_pct, control$post)
  control_ts <- zoo(control_sel, dates)
  
  # Analysis 10.a: DeltaCoverage (t) ~ DeltaCoverage (t-1) + DeltaRTSI (t-1)
  analysis_10 <-
    dyn$lm(control.post ~ lag(control.post,-1) + lag(control.rtsi_pct,-1), data = control_ts)
  
  # Analysis 11.a: DeltaCoverage (t) ~ DeltaCoverage (t-2) + DeltaRTSI (t-2)
  analysis_11 <-
    dyn$lm(
      control.post ~ lag(control.post,-1) + lag(control.rtsi_pct,-1) + lag(control.post, -2) + lag(control.rtsi_pct, -2),
      data = control_ts
    )

  # Plot and inspect results
  print(elem)

   print(summary(analysis_10))
  print(summary(analysis_11))
  #print(summary(analysis_e3))
  
  #plot(analysis_10)
  plot(analysis_11)
  
  # Save results
  analysis_10_c[[elem]] <- analysis_10
  analysis_11_c[[elem]] <- analysis_11
}
```
